---
title: "Malnutrition and Health Indicators Across Time and Population"
Name: Hitesh Karnatak
Student Number: 22263508
Lecturer: Damien Dupre
Assignment2: Quarto Report
Date: 27-04-2023
Subject: Data Analytics & Visualisation (MT5000)
format: html
execute:
  echo: false
  warning: false
  message: false
---

**Indicators of Nutritional Status in Five Countries in 2020**

```{r}
library(tidyverse)
library(ggplot2)
library(maps)

# Load data
gapminder <- read.csv("world_data.csv", stringsAsFactors = FALSE)

# Filter data for 2007
library(dplyr)
library(readr)

gapminder <- gapminder %>%
  mutate(Year = as.character(Year),
         year = parse_number(Year))

world_data_2020<-gapminder
#world_data_2020 <- world_data_2020 %>%
  #left_join(world_data, by = "Country")

# Create new data frame with only countries with numeric population values
world_data_2020_clean <- world_data_2020 %>%
  filter(!is.na(population) & is.numeric(population))

world_data_2020_clean$populationmill<-world_data_2020_clean$population/1000000
# View the new data frame
#View(world_data_2020_clean)

# Load map data
map_world <- map_data("world")

# Join population data with map data
map_pop_2020 <- merge(x = map_world, y = world_data_2020_clean, by.x = c("region") ,by.y =  "Country",all.x = T)

# Create choropleth map using ggplot2
ggplot(data = map_pop_2020) +
  aes(x = long, y = lat, group = group, fill = populationmill) +
  geom_polygon() +
  scale_fill_gradient(low = "orange", high = "red", na.value = "grey") +
  labs(
    title = "Child Malnutrition in 2020: A Global Overview of Stunting and Wasting of top 5 Country",
    subtitle = "",
    caption = "",
    x = "Longitude",
    y = "Latitude",
    fill = "Country Population in Million"
  ) +
  theme_bw()
```

The populations and height-for-age and weight-for-height metrics for five nations in 2020. Children in Bangladesh, India, and Pakistan have poor height-for-age and weight-for-height ratios, indicating a large burden of malnutrition. Malnutrition burdens are lower in Brazil and China. 

**Life Expectancy and Population in Five Countries in 2020**

```{r}
library(ggplot2)
library(tidyverse)
data <- read.csv('data_2020.csv')
data$Avg..Population..total<-data$Avg..Population..total/1000000
#data$Avg..Population..total<-as.numeric(data$Avg..Population..total)
x <- data$Avg..Population..total
y <- data$Avg..Life.expectancy.at.birth..total..years.

model <- lm(y ~ x)

p <- ggplot(data , aes(x = x, y = y)) +
  geom_point(color = 'blue') +
  geom_smooth(method = 'lm', color = 'red', formula = y ~ x, se = FALSE) +
  ggtitle("Life Expectancy and Population in Five Countries in 2020") +
  xlab("Population in Million") +
  ylab("Life expectancy at birth, total (years)") +
  theme_bw() + 
  geom_text(aes(label = Country..Unicef.Metadata.Csv.), color = 'black', hjust = -0.2, vjust = 0.2, size = 3)

line_eqn <- paste0("y = ", round(coef(model)[2], 2), "x + ", round(coef(model)[1], 2))
p <- p + annotate("text", x = max(x), y = min(y), label = line_eqn, hjust = 1, vjust = -0.2, size = 3.5)

print(p)

```


The population and life expectancy of five nations in 2020: Hong Kong, Japan, North Korea, Macao, and Singapore. The average life expectancy of Japan, Hong Kong, and Macao was 85 years, whereas Singapore had a slightly lower life expectancy of 84 years. With a life expectancy of 73 years, North Korea was the lowest. Japan had the most people, while Singapore and Macao had the fewest. 

**Male and Female Prevalence of Low Height-for-Age and Weight-for-Height**

```{r}
library(tidyverse)

data_2_2020 <- read.csv("unicef_indicator_2.csv")
data_2_2020 <- data_2_2020[data_2_2020$time_period == "2020", ]

#head(data_2_2020)

data_2_2020 <- data_2_2020[order(-data_2_2020$obs_value), ]

#head(data_2_2020)

top_5 <- head(data_2_2020, 5)

values <- top_5 %>%
  group_by(sex) %>%
  summarize(avg_value = mean(obs_value))

ggplot(values, aes(x = reorder(sex, avg_value), y = avg_value, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.5) +
  scale_fill_manual(values = c("red", "green", "blue")) +
  labs(x = "Sex", y = "Avg. Observed Value", title = "Observed value based on sex") +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1)) +
  coord_flip()

```

The frequency of low height-for-age and low weight-for-height in girls and boys with a -2SD cutoff. Males have a larger prevalence of low values (2.356%) than females (1.328%), for a total prevalence of 1.828% in both sexes. Malnutrition is indicated by low height-for-age and weight-for-height, which can result in unfavorable health consequences and impaired cognitive development. 

**Malnutrition Rates in Children from 1986 to 2021**

```{r}
df <- as.data.frame(read.csv("file.csv", header = TRUE, row.names = 1))

df <- read.csv("file.csv", header = TRUE, row.names = 1)
df <- t(df)
df_f <- df[-c(1:3), ]

df_f <- as.data.frame(df[-c(1:3), ])

df_f <- subset(df_f, complete.cases(df_f))

colnames(df_f) <- c("X__1", "X__2", "X__3")
df_f$X__1<-as.integer(df_f$X__1)

df_f$X__2<-round(as.numeric(df_f$X__2),1)
df_f$X__3<-round(as.numeric(df_f$X__3),1)
ggplot(data = df_f, aes(x = X__1, y = X__2, group = 1)) +
  geom_line(aes(color = "Female"), linewidth = 0.5) +
  geom_line(aes(y = X__3, color = "Male"), linewidth = 0.5) +
  xlab("Time Period") +
  ylab("Std. Dev. of Obs Value") +
  ggtitle("Prevalence of Malnutrition Among Males and Females") +
  theme(axis.line = element_line(color = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) +
  scale_color_manual(values = c("Female" = "red", "Male" = "blue")) 
```

From 1986 to 2021, the table displays the percentage of girls and men with a height-for-age less than -2 standard deviations and a weight-for-height less than -2 SD. This disorder affected 61.1% of females and 58.6% of men in 1988, but only 1.365% of females and 1.699% of males in 2020.

